Open-source landscape for 3D CSEM modelling

### 2020-09-08 manuscript/Manuscript.tex
Working on manuscript, EOD 08 Sep

%%fakesection ===    ABSTRACT    ===
\begin{abstract}
%
Large-scale modelling of three-dimensional (3D) controlled-source electromagnetic (CSEM) responses used to be feasible only for large companies and research consortia. However, this changed over the last few years, and today there exists a selection of different open-source codes which are up to the task.
%
Four different open-source 3D CSEM codes are considered in this work, where in all cases at least the user-facing functionality is written in Python. Two codes, \simpeg and \emg3d, use finite volumes, the other two codes, \custem and \petgem, use finite elements. Three different mesh-types are used, simple tensor meshes, unstructured tetrahedral meshes, and octree meshes. Two models are used to compare the codes, a layered background containing three resistive blocks, and Marlim R3D, a complex, real-world, open-source resistivity data.
%
The comparison shows that all four codes are able to compute the model to sufficient precision. The biggest difference between the codes are actually the used mesh types. The required runtime and memory varies greatly between the codes. However, these requirements depend mainly on the underlying solver, and not on the actual CSEM codes.
%
The collaboration of four code maintainer trying to achieve the same task brought in the end all four codes a significant step further. We hope that these results may be useful for the entire CSEM community, and we invite and encourage the community to make more code, modelling scripts, and results publicly available. For a start, all the scripts to reproduce our shown results are available.
%
\end{abstract}

@@

In the 1990's, computer became sufficiently powerful that 3D modelling gained traction, which resulted amongst other publications in the book \emph{Three-Dimensional Electromagnetics} by the SEG \citep{B.SEG.99.Oristaglio}. Often cited publications from that time are \cite{RSC.94.Mackie}, 3D MT computation; \cite{RS.94.Druskin}, frequency- and time-domain modelling using a Yee grid and a global Krylov subspace approximation; and \cite{RS.96.Alumbaugh, GJI.97.Newman}, low-to-high frequency computation on massively parallel computers.

The continuous improvement of computing power and the CSEM boom in the early 2000's in the hydrocarbon industry led to a wealth of developed numerical solutions and according publications. The most common applied methods to solve Maxwell's equation are the IE method \citep{GJI.74.Raiche, RS.02.Hursan, GEO.06.Zhdanov, GP.10.Tehrani, CAG.16.Kruglyakov, MGS.17.Kruglyakov} and different variations of the differential equation (DE) method, such as finite differences (FD) \citep{IEEE.66.Yee, GEO.93.Wang, RSC.94.Mackie, RS.94.Druskin, GEO.09.Streich, CAG.13.Sommer}, finite elements (FE) \citep{GEO.04.Commer, GJI.11.Schwarzbach,GEO.12.daSilva, GJI.13.Grayver, GJI.13.Puzyrev, SEG.16.Zhang}, and finite volumes (FV) \citep{EM.90.Madsen, SIAM.01.Haber, PIER.01.Clemens, GEO.14.Jahandari}. There are also many different types of discretisation, where the most common ones are regular grids (Cartesian, rectilinear), mostly using a Yee grid \citep{IEEE.66.Yee} or a Lebedev grid \citep{CMMP.64.Lebedev}, but also unstructured tetrahedral grids \citep{SEG.16.Zhang, CAG.17.Cai}, hexagonal meshes \citep{CAG.14.Cai}, or OcTree meshes \citep{ECP.07.Haber}. Very well written overviews about the different approaches to 3D EM modelling are given by \cite{SG.05.Avdeev} and \cite{SG.10.Borner}.

In the last 15 years, the publications with regards to 3D EM modelling grew tremendously, driven by the ever increasing computing power. One reason why there are so many reports about this topic results from the variety of solution techniques for the systems of linear equations (SLE). They can be distinguished between direct solvers \citep{GEO.09.Streich, GP.14.Chung, GEO.14.Jaysaval, GEO.15.Grayver, SEG.15.Oh, GJI.18.Wang}, indirect solvers \citep{GP.06.Mulder, GJI.15.Jaysaval}, or a combination of both, so-called hybrid solvers \citep{GEO.18.Liu}. The solvers often use preconditioners such as the multigrid method \citep{SIAM.02.Aruliah, GP.06.Mulder, GJI.16.Jaysaval}.

Many of the advancements made in the EM modelling community over the past decades have required that authors develop new implementation from scratch. These codes often provided the research group or company with a competitive advantage for a time, and thus the source codes were often kept internal. In some cases, executables have been made available for academic purposes upon request or to sponsoring consortium members. As the field continues to mature, advancements become more incremental, and particularly in an applied field, many advancements are driven by new use-cases and applications that were not considered by the original authors. In his review on EM modelling and inversion \cite{SG.05.Avdeev} concludes with the following statement: «\emph{The most important challenge that faces the EM community today is to convince software developers to put their 3-D EM forward and inverse solutions into the public domain, at least after some time. This would have a strong impact on the whole subject and the developers would benefit from feedback regarding the real needs of the end-users.}» Further, \cite{EXG.19.Oldenburg} argue that an open-source paradigm has the potential to accelerate multidisciplinary research by encouraging the development of modular, interoperable tools that are supported by a community of researchers.

In many domains of science, it is becoming more common for researchers to release source-code with an open license that allows use and modification (e.g., see the Open Source Initiative approved licenses: \href{https://opensource.org/licenses}{opensource.org/licenses}). This is an important step for improving reproducibility of research, «\emph{to provide the means by which the reader can verify the validity of the results and make use of them in further research}» \citep{GEO.17.Broggini}. Going a step beyond releasing open-source software, many groups adopt an open model of development where code is hosted and versioned in an online repository, all changes are public, and users can engage by submitting and track issues. Community oriented projects further engage by encouraging pull-requests, which are suggested changes to the code. Successful, well-maintained projects often have unit-testing and continuous integration that runs those tests with any changes to the code. Additionally, documentation that includes examples and tutorials is an important component for on-boarding new users and contributors. As a result, in many areas of the geosciences, we are seeing a shift away from a one-way distribution of code towards building global communities around open projects. Related projects within the realm of geophysical exploration are, e.g., \texttt{pyGIMLi} \citep{CAG.17.Rucker} or \texttt{Fatiando} \citep{JOSS.18.Uieda}
% Others coming to mind?
% DW: - https://github.com/fancompute/ceviche
% RR: - no thoughts about geophysics, but there are many projects for
%       related topics such as gmt, gis, mesh-gen
% LH  - https://github.com/softwareunderground/awesome-open-geoscience ?
%     - Madagascar in seismic?

In this paper, we discuss the topic of open-source developments with respect to 3D CSEM modelling. Even though it comprises still the minority of developments, the landscape in this field of research changed in the last five years. Previously, there were only closed-source codes owned by companies or consortia and individual projects, often without proper documentation, that could be requested from the authors. We introduce and compare four recent open-source projects with focus on 3D CSEM modelling. All presented codes are in the Python ecosystem and use either the FV method on structured grids or the FE method on unstructured tetrahedral meshes.

Furthermore, we deal with the topic of verification. \cite{GJI.13.Miensopust} presents a review of two workshops dealing with the verification of magnetotelluric forward and inversion codes, but we are not aware of any comparable study or benchmark suite for CSEM data. Hence, we first conduct a similar exercise for CSEM codes. We want our efforts to help to ensure the reliability of simulations, not only carried out with the available open-source codes, but also for individual closed-source projects. Analytical and semi-analytical solutions only exist for simple halfspace or layered-Earth models, which served mostly to verify new codes. Beyond these simple models, the only objective possibility to ensure the accuracy of solutions is by comparing results from different modellers. If different discretisations and implementations of Maxwell's equations yield the same result, it gives confidence in their accuracy.

We simulate EM fields for a layered background model with vertical transverse isotropy, containing three blocks, as well as for the complex marine Marlim R3D model. These two model designs as well as the corresponding results from four different codes provide a benchmark for new codes to be compared to and validated with. In the following section, we introduce the codes under consideration. Afterwards, we present the considered benchmark cases in detail and present the modelling results of our four codes in terms of accuracy and computational performance. Beyond that, we extensively discuss important points that control the performance and suitability of our FV and FE codes, including mainly considerations about the mesh design and the choice of solvers. We conclude with a summary of our results and a motivation for the EM community to not only continue to extend the landscape of open-source codes but to also create a landscape of open-source benchmark models.

\itodo{Introduce Maxwell's Equations or not? Miensopust et al., 2013, don't.}
% Following the conventions of Ward and Hohmann \cite{Ward1988}, the quasi-static Maxwell's equations are

@@

The four codes under consideration are, in alphabetical order, \custem \citep{GEO.19.Rochlitz}, \emg3d \citep{JOSS.19.Werthmuller}, \petgem \citep{CAG.18.CastilloReyes, GJI.19.CastilloReyes}, and \simpeg \citep{CAG.15.Cockett, CAG.17.Heagy}. All four codes have their user-facing routines written in Python, and all of them make heavy use of \texttt{NumPy} \citep{CSE.11.VanDerWalt} and \texttt{SciPy} \citep{NM.20.Virtanen}. The four of them are “modern” open-source projects, meaning that they come with both, an open-source license and an online-hosted version-control system with tracking possibilities (raising issues, filing pull requests). All developments comprise an extensive online documentation with many examples and have continuous integration to some degree. Newer package-management systems such as Conda, Docker, or pip significantly supported us to make our codes easily available and installable on any of the major operating system without any user-side compilations. Each one can be downloaded and installed with a single command.

In the following, we provide a quick overview of the codes. It is, however, beyond the scope of this article to go into every detail of the different modellers, and we refer to their documentations for more details. An overview comparison of the different codes is given in Table~\ref{tbl:codecomparison}. All codes have in common that they solve the weak formulation of Maxwell's equation in its differential form (DE) under the quasistatic or diffusive approximation, hence neglecting displacement currents, which is a common approximation for the low frequency range commonly applied in CSEM. The machines on which the different codes were run are listed in Table~\ref{tbl:machines} together with the responsible operator.
% LH: SimPEG also interfaces to the scipy solvers; but in practice those are
%     use much less than Pardiso / mumps
% => I leave them out for now, as you did not use them for this paper either,
%    right? Maybe to mention in the SimPEG-section.

\tabl[btp]{codecomparison}{Comparison of the four codes under consideration.
  Note that \emg3d is a solver on its own, while the other codes implement
  third-package solvers such as \texttt{PETSc} \citep{Preprint.Abhyankar},
  \texttt{MUMPS} \citep{SIAM.01.Amestoy}, or \texttt{PARDISO}
  \citep{FGCS.04.Schenk}.
}{
  \centering
  \footnotesize
\begin{tabularx}{\linewidth}{lXXXX}
  \toprule
  %
  & \custem & \emg3d & \petgem & \simpeg  \\
  \midrule
  %
  Home & \href{https://custem.rtfd.io}{custem.rtfd.io}
       & \href{https://empymod.github.io}{empymod.github.io}
       & \href{http://petgem.bsc.es}{petgem.bsc.es}
       & \href{https://docs.simpeg.xyz}{simpeg.xyz} \\
  License & GPL-3.0 & Apache-2.0 & GPL-3.0 & MIT \\
  Installation & \texttt{conda}
               & \texttt{pip}; \texttt{conda}
               & \texttt{pip}
               & \texttt{pip}; \texttt{conda} \\
  Comp. Dom. & frequency \& time & frequency & frequency & frequency \& time \\
  Method & FE & FV & FE & FV \\
  Mesh & tetrahedral & rectilinear & tetrahedral & recti- \& curvilinear \\
  BC & PMC; PEC & PEC & PEC & PEC; PMC \\
  Solver & \texttt{MUMPS} & \emg3d & \texttt{PETSc}; \texttt{MUMPS} &
           \texttt{PARDISO}; \texttt{MUMPS} \\
  %
  \bottomrule
\end{tabularx}}%
%

%
\tabl[btp]{machines}{List of computer and operating system (hardware and
software) on which the different codes were run, together with the operator.}{
  \centering
  \footnotesize
  \begin{tabularx}{\linewidth}{lXl}
  \toprule
  %
  Code & Computer and Operating System & Operator \\
  \midrule
  %
  \custem & \textbf{PowerEdge R940 server}; 144 Xeon Gold 6154 CPU @2.666 GHz;
            ~3 TB DDR4 RAM; Ubuntu 18.04
          & Raphael Rochlitz \\
  \emg3d  & \textbf{Laptop}; i7-6600U CPU@2.6 GHz x4; 16 GB of memory,
            Ubuntu 18.04 & Dieter Werthmüller \\
  \petgem & \textbf{Marenostrum4 (supercomputer)}. Intel Xeon Platinum from
            Skylake generation; 2 sockets Intel Xeon Platinum 8160 CPU with 24
            cores each @2.10GHz for a total of 48 cores per node; 386 GB DDR4
            RAM per node; SuSE Linux Enterprise
          & Octavio Castillo-Reyes \\
          \simpeg & \itodo{Todo Lindsey}
          & Lindsey Heagy \\
  %
  \bottomrule
\end{tabularx}}%
%

@@

The customisable electromagnetic modelling Python toolbox custEM was developed for simulating arbitrary complex 3D CSEM geometries with a focus on semi-airborne setups, but it supports also land-based, airborne, coastal, or marine environments. Multiple electric or magnetic field or potential finite-element approaches were implemented as total or secondary field formulations. The finite-element kernel, including higher order basis functions and parallelisation, relies on the FEniCS project \citep{B.SPR.12.Logg, B.SPR.16.Langtangen}. The resulting systems of linear equations are solved with MUMPS \citep{SIAM.01.Amestoy}, which is a very robust but memory consuming choice. Primary field solutions are supplied by the COMET \citep{GEO.20.Skibbe} package.

The toolbox considers generally anisotropic petrophysical properties. Even though changes of the conductivity is mainly of interest for CSEM modelling, the electric permittivity and magnetic permeability can be taken into account using the preferred electric field approach on Nédélec elements. Recently, induced polarization parameters in frequency-domain and three methods for simulating time-domain responses were added to custEM. The provided meshing tools based on TetGen \citep{TOM.15.Si} and functionalities of pyGIMLi \citep{CAG.17.Rucker} facilitate the generation of tetrahedral meshes including layered-earth geometries with topography or bathymetry and anomalous bodies which are allowed to be connected or to reach the surface.

@@

The 3D CSEM code \emg3d is a multigrid \citep{CMMP.64.Fedorenko} solver for 3D electromagnetic diffusion following \cite{GP.06.Mulder}, with tri-axial electrical anisotropy, isotropic electric permittivity, and isotropic magnetic permeability. The matrix-free solver can be used as main solver or as preconditioner for one of the Krylov subspace methods implemented in SciPy, and the governing equations are discretized on a staggered grid by the finite-integration technique \citep{AEU.77.Weiland}, which is a finite-volume generalization of a Yee grid. The code is written completely in Python using the NumPy and SciPy stacks, where the most time- and memory-consuming parts are sped up through jitted (just-in-time compiled) functions using Numba \citep{LLVM.15.Lam}. The multigrid method is characterized by almost linear scaling both in terms of runtime (CPU) and memory (RAM) usage, and it is therefore a comparably low-memory consumption solver.

@@

Current development is focused on adding basic inversion capabilities, and a further plan is to implement a hook in order to use \emg3d as a frequency-domain CSEM solver within the grander \simpeg framework.

@@

\petgem permits to locate the source and receivers anywhere in the computational domain (e.g., sediments, seafloor, sea, ground), allowing to analyze the physical environment of the electric responses and how parameters impact them (e.g., frequency, conductivity, dependence on mesh setup, basis order, solver type, among others). Furthermore, \petgem implements a semi-adaptive mesh strategy ($hp$ mesh refinement) based on physical parameters and on polynomial order to satisfy quality criteria chosen by the user. Nonetheless, only Horizontal Electric Dipole (HED) has been implemented.

For the parallel forward modelling computations, a highly scalable MPI domain decomposition allows reducing run-times and the solution of large-scale modelling cases. This strategy is capable of exploiting the parallelism offered by both modest multi-core computers and cutting-edge clusters (e.g., High-Performance Computing architectures).

@@

\simpeg is a modular toolbox for simulations and gradient based inversions in geophysics. Current functionality includes modelling and inversion capabilities for gravity, magnetics, direct current resistivity, induced polarization, electromagnetics (time, frequency, as well as natural sources such as magnetotellurics), and fluid flow (Richards equation). It is a community driven project that aims to support researchers by providing a flexible, extensible framework and to facilitate integration of data sets from a variety of geophysical methods, including joint inversions \citep{GJI.20.Astic}, by providing a common interface to each method. \imycom[DW]{Lindsey: «support researchers» downplays the fact that \simpeg is used by various companies as well\dots}

@@

The electromagnetics implementations use a staggered grid finite volume approach where physical properties are discretized at cell centers, fields on cell edges and fluxes on cell faces \citep{IEEE.66.Yee}.
%(for an overview of discretizing Maxwell's equations, we recommend \cite{B.SIAM.14.Haber}).
There are two different discretization strategies implemented for Maxwell's equations: (a) the EB-formulation, which discretizes the electric field ($\vec{e}$) on cell edges and the magnetic flux density ($\vec{b}$) on cell faces, and (b) the HJ-formulation, which discretizes the magnetic field ($\vec{h}$) on cell edges and the current density ($\vec{j}$) on cell faces. In both cases the physical properties: electrical conductivity / resistivity and magnetic permeability are discretized at cell centers. Having multiple implementations allows for testing that compares results from each approach, as well as the representation of both electrical and magnetic sources on cylindrically symmetric meshes. \simpeg supports variable magnetic permeability and full-tensor anisotropy for the physical properties.

\simpeg interfaces to various solvers including \texttt{PARDISO} \citep{FGCS.04.Schenk}, \texttt{MUMPS} \citep{SIAM.01.Amestoy}, as well as those implemented in the \texttt{scipy.sparse.linalg} module \citep{NM.20.Virtanen}. For problems with many sources (e.g. airborne EM), \simpeg contains machinery for domain-decomposition approaches; see for example \cite{GEO.20.Fournier}.

\section{Numerical Verifications}

@@

@@

\section{Acknowledgment}

>>>>

\begin{abstract}
%
Large-scale modelling of three-dimensional (3D) controlled-source electromagnetic (CSEM) responses used to be feasible only for large companies and research consortia. However, this changed over the last few years, and today there exists a selection of different open-source codes which are up to the task.
%
We have a closer look at four different open-source 3D CSEM codes, where in all cases at least the user-facing functionality is written in Python. Two of the codes use finite volumes and the other two use finite elements. For the modelling we consider three different mesh-types, simple tensor meshes, unstructured tetrahedral meshes, and octree meshes. Two models are used to compare the codes, a layered background containing three resistive blocks, and Marlim R3D, a complex, real-world, open-source resistivity data.
%
The comparison shows that all four codes are able to compute the different models to sufficient precision. The biggest difference between the codes are actually the used discretization. The required runtime and memory varies greatly between the codes. However, these requirements depend mainly on the underlying solver, and not on the actual CSEM codes.
%
The collaboration of four code maintainer trying to achieve the same task brought in the end all four codes a significant step further. We hope that these results may be useful for the entire CSEM community, and we invite and encourage the community to make more code, modelling scripts, and results publicly available. For a start, all the scripts to reproduce our shown results are available.
%
\end{abstract}

\itodo{We probably should be careful to not mention the actual code names too often. I also removed them from the abstract.}

\itodo{Notes on orthography: I settled on discretize with a "z". My dictionary says that discretize ("z") is AE and discretise ("s") is BE, but that "z" is also used in BE in mathematical contexts. Is that fine or any other opinions?}

@@

In the 1990's, computer became sufficiently powerful that 3D modelling gained traction, which resulted amongst other publications in the book \emph{Three-Dimensional Electromagnetics} by the SEG \citep{B.SEG.99.Oristaglio}. Often cited publications from that time are \cite{RSC.94.Mackie} for 3D MT computation; \cite{RS.94.Druskin} for frequency- and time-domain modelling using a Yee grid and a global Krylov subspace approximation; and \cite{RS.96.Alumbaugh, GJI.97.Newman} for low-to-high frequency computation on massively parallel computers.

The continuous improvement of computing power and the CSEM boom in the early 2000's in the hydrocarbon industry led to a wealth of developed numerical solutions and according publications. The most common applied methods to solve Maxwell's equation are the IE method \citep{GJI.74.Raiche, RS.02.Hursan, GEO.06.Zhdanov, GP.10.Tehrani, CAG.16.Kruglyakov, MGS.17.Kruglyakov} and different variations of the differential equation (DE) method, such as finite differences (FD) \citep{IEEE.66.Yee, GEO.93.Wang, RSC.94.Mackie, RS.94.Druskin, GEO.09.Streich, CAG.13.Sommer}, finite elements (FE) \citep{GEO.04.Commer, GJI.11.Schwarzbach,GEO.12.daSilva, GJI.13.Grayver, GJI.13.Puzyrev, SEG.16.Zhang}, and finite volumes (FV) \citep{EM.90.Madsen, SIAM.01.Haber, PIER.01.Clemens, GEO.14.Jahandari}. There are also many different types of discretization, where the most common ones are regular grids (Cartesian, rectilinear), mostly using a Yee grid \citep{IEEE.66.Yee} or a Lebedev grid \citep{CMMP.64.Lebedev}, but also unstructured tetrahedral grids \citep{SEG.16.Zhang, CAG.17.Cai}, hexagonal meshes \citep{CAG.14.Cai}, or OcTree meshes \citep{ECP.07.Haber}. Very well written overviews about the different approaches to 3D EM modelling are given by \cite{SG.05.Avdeev} and \cite{SG.10.Borner}.

In the last 15 years, the publications with regards to 3D EM modelling grew tremendously, driven by the ever increasing computing power. One reason why there are so many publications about this topic results from the variety of solution techniques for the systems of linear equations (SLE). They can be distinguished between direct solvers \citep{GEO.09.Streich, GP.14.Chung, GEO.14.Jaysaval, GEO.15.Grayver, SEG.15.Oh, GJI.18.Wang}, indirect solvers \citep{GP.06.Mulder, GJI.15.Jaysaval}, or a combination of both, so-called hybrid solvers \citep{GEO.18.Liu}. The solvers often use preconditioners such as the multigrid method \citep{SIAM.02.Aruliah, GP.06.Mulder, GJI.16.Jaysaval}.

Many of the advancements made in the EM modelling community over the past decades have required that authors develop new implementations from scratch. These codes often provided the research group or company with a competitive advantage for a time, and thus the source codes were often kept internal. In some cases, executables have been made available for academic purposes upon request or to sponsoring consortium members. As the field continues to mature, advancements become more incremental, and particularly in an applied field, many advancements are driven by new use-cases and applications that were not considered by the original authors. In his review on EM modelling and inversion \cite{SG.05.Avdeev} concludes with the following statement: «\emph{The most important challenge that faces the EM community today is to convince software developers to put their 3-D EM forward and inverse solutions into the public domain, at least after some time. This would have a strong impact on the whole subject and the developers would benefit from feedback regarding the real needs of the end-users.}» Further, \cite{EXG.19.Oldenburg} argue that an open-source paradigm has the potential to accelerate multidisciplinary research by encouraging the development of modular, interoperable tools that are supported by a community of researchers.

In many domains of science, it is becoming more common for researchers to release source-code with an open license that allows use and modification (e.g., see the Open Source Initiative approved licenses: \href{https://opensource.org/licenses}{opensource.org/licenses}). This is an important step for improving reproducibility of research, «\emph{to provide the means by which the reader can verify the validity of the results and make use of them in further research}» \citep{GEO.17.Broggini}. Going a step beyond releasing open-source software, many groups adopt an \emph{open model} of development where code is hosted and versioned in an online repository, all changes are public, and users can engage by submitting and track issues. Community oriented projects further engage by encouraging pull-requests, which are suggested changes to the code. Successful, well-maintained projects often have unit-testing and continuous integration that runs those tests with any changes to the code. Additionally, documentation that includes examples and tutorials is an important component for on-boarding new users and contributors. As a result, in many areas of the geosciences, we are seeing a shift away from a one-way distribution of code towards building global communities around open projects. Notable other open-source Python projects within the same realm are \texttt{pyGIMLi} \citep{CAG.17.Rucker}, \texttt{Fatiando} \citep{JOSS.18.Uieda}, \texttt{GemPy} \citep{GMD.19.DeLaVarga}, and \texttt{PyVista} \citep{JOSS.19.Sullivan}.

In this paper, we discuss the topic of open-source developments with respect to 3D CSEM modelling. Even though it comprises still the minority of developments, the landscape in this field of research changed in the last five years. Previously, there were only closed-source codes owned by companies or consortia and individual projects, that were only available directly from the authors upon request. These codes often came without proper documentation, and also often without a proper License, which essentially means that they are not open source. We introduce and compare four recent open-source projects with a focus on 3D CSEM modelling. All presented codes are in the Python ecosystem and use either the FV method on structured grids or the FE method on unstructured tetrahedral meshes.

Furthermore, we deal with the topic of verification. Analytical and semi-analytical solutions only exist for simple halfspace or layered-Earth models, which served mostly to verify new codes. Beyond these simple models, the only objective possibility to ensure the accuracy of solutions is by comparing results from different modellers. If different discretizations and implementations of Maxwell's equations yield the same result, it gives confidence in their accuracy. \cite{GJI.13.Miensopust} presents a review of two workshops dealing with the verification of magnetotelluric forward and inversion codes, but we are not aware of any comparable study or benchmark suite for CSEM data. Hence, we first conduct a similar exercise for CSEM codes. We want our efforts to help to ensure the reliability of simulations, not only carried out with the available open-source codes, but also for individual closed-source projects.

We simulate EM fields for a layered background model with vertical transverse isotropy, containing three blocks, as well as for the complex marine Marlim R3D model. These two model designs as well as the corresponding results from four different codes provide a benchmark for new codes to be compared to and validated with. In the following section, we introduce the codes under consideration. Afterwards, we present the considered benchmark cases in detail and present the modelling results of our four codes in terms of accuracy and computational performance. Beyond that, we extensively discuss important points that control the performance and suitability of our FV and FE codes, including considerations about the mesh design and the choice of solvers. We conclude with a summary of our results and a motivation for the EM community to not only continue to extend the landscape of open-source codes but to also create a landscape of open-source benchmark models.

\itodo{Introduce Maxwell's Equations or not? Miensopust et al., 2013, do not touch on any equation. We agreed that we wait and see how large the paper becomes, and see at the end if it should be added or not.}
% Following the conventions of Ward and Hohmann \cite{Ward1988}, the quasi-static Maxwell's equations are

@@

The four codes under consideration are, in alphabetical order, \custem \citep{GEO.19.Rochlitz}, \emg3d \citep{JOSS.19.Werthmuller}, \petgem \citep{CAG.18.CastilloReyes, GJI.19.CastilloReyes}, and \simpeg \citep{CAG.15.Cockett, CAG.17.Heagy}. All four codes have their user-facing routines written in Python, and all of them make heavy use of \texttt{NumPy} \citep{CSE.11.VanDerWalt} and \texttt{SciPy} \citep{NM.20.Virtanen}. The four of them are \mycom[DW]{change to open dev} “modern” open-source projects, meaning that they come with both, an open-source license and an online-hosted version-control system with tracking possibilities (raising issues, filing pull requests). All developments comprise an extensive online documentation with many examples and have continuous integration to some degree. Newer package-management systems such as Conda, Docker, or pip significantly supported us to make our codes easily available and installable on any of the major operating system without any user-side compilations. Each one can be downloaded and installed with a single command.

In the following, we provide a quick overview of the codes. It is, however, beyond the scope of this article to go into every detail of the different modellers, and we refer to their documentations for more details. An overview comparison of the different codes is given in Table~\ref{tbl:codecomparison}. All codes have in common that they solve the weak formulation of Maxwell's equation in its differential form (DE) under the quasi-static or diffusive approximation, hence neglecting displacement currents, which is a common approximation for the low frequency range commonly applied in CSEM. The machines on which the different codes were run are listed in Table~\ref{tbl:machines} together with the responsible operator.
% LH: SimPEG also interfaces to the scipy solvers; but in practice those are
%     use much less than Pardiso / mumps
% => I leave them out for now, as you did not use them for this paper either,
%    right? Maybe to mention in the SimPEG-section.

\tabl[btp]{codecomparison}{Comparison of the four codes under consideration. Note that \emg3d is a solver on its own, while the other codes implement third-package solvers such as \texttt{PETSc} \citep{Preprint.Abhyankar}, \texttt{MUMPS} \citep{SIAM.01.Amestoy}, or \texttt{PARDISO} \citep{FGCS.04.Schenk}.}{
  \centering
  \footnotesize
\begin{tabularx}{\linewidth}{lXXXX}
  \toprule
  %
  & \custem & \emg3d & \petgem & \simpeg  \\
  \midrule
  %
  Home & \href{https://custem.rtfd.io}{custem.rtfd.io}
       & \href{https://empymod.github.io}{empymod.github.io}
       & \href{http://petgem.bsc.es}{petgem.bsc.es}
       & \href{https://docs.simpeg.xyz}{simpeg.xyz} \\
  License & GPL-3.0 & Apache-2.0 & GPL-3.0 & MIT \\
  Installation & \texttt{conda}
               & \texttt{pip}; \texttt{conda}
               & \texttt{pip}
               & \texttt{pip}; \texttt{conda} \\
  Comp. Dom. & frequency \& time & frequency & frequency & frequency \& time \\
  Method & FE & FV & FE & FV \\
  Mesh & tetrahedral & rectilinear & tetrahedral & recti- \& curvilinear \\
  BC & PMC; PEC & PEC & PEC & PEC; PMC \\
  Solver & \texttt{MUMPS} & \emg3d & \texttt{PETSc}; \texttt{MUMPS} &
           \texttt{PARDISO}; \texttt{MUMPS} \\
  %
  \bottomrule
\end{tabularx}}%
%

%
\tabl[btp]{machines}{List of computer and operating system (hardware and
software) on which the different codes were run, together with the operator.}{
  \centering
  \footnotesize
  \begin{tabularx}{\linewidth}{lXl}
  \toprule
  %
  Code & Computer and Operating System & Operator \\
  \midrule
  %
  \custem & \textbf{PowerEdge R940 (server)}; 144 Xeon Gold 6154 CPU @2.666 GHz;
            ~3 TB DDR4 RAM; Ubuntu 18.04
          & Raphael Rochlitz \\
  \emg3d  & \textbf{Dell Latitude (laptop)}; i7-6600U CPU@2.6 GHz x4; 15.5 GiB of memory,
            Ubuntu 18.04 & Dieter Werthmüller \\
  \petgem & \textbf{Marenostrum4 (supercomputer)}. Intel Xeon Platinum from
            Skylake generation; 2 sockets Intel Xeon Platinum 8160 CPU with 24
            cores each @2.10GHz for a total of 48 cores per node; 386 GB DDR4
            RAM per node; SuSE Linux Enterprise
          & Octavio Castillo-Reyes \\
  \simpeg & \textbf{GKE n2-custom (Google cloud)}. Intel Cascade Lake; 8 vCPUs, 355 GB memory; Ubuntu 16.04
          & Lindsey Heagy \\
  %
  \bottomrule
\end{tabularx}}%
%
\mycom[DW]{Introduce dof, PEC, PMC}

@@

The customisable electromagnetic modelling Python toolbox custEM was developed for simulating arbitrary complex 3D CSEM geometries with a focus on semi-airborne setups, but it supports also land-based, airborne, coastal, or marine environments. Multiple electric or magnetic field or potential finite-element approaches were implemented as total or secondary field formulations. The finite-element kernel, including higher order basis functions and parallelisation, relies on the FEniCS project \citep{B.SPR.12.Logg, B.SPR.16.Langtangen}. The resulting systems of linear equations are solved with MUMPS \citep{SIAM.01.Amestoy}, which is a very robust but memory consuming choice. Primary field solutions are supplied by the \texttt{COMET} package \citep{GEO.20.Skibbe}.

The toolbox considers generally anisotropic petrophysical properties. Even though changes of the conductivity is mainly of interest for CSEM modelling, the electric permittivity and magnetic permeability can be taken into account using the preferred electric field approach on Nédélec elements. Recently, induced polarisation parameters in frequency-domain and three methods for simulating time-domain responses were added to custEM. The provided meshing tools are based on TetGen \citep{TOM.15.Si} and functionalities of pyGIMLi \citep{CAG.17.Rucker} facilitate the generation of tetrahedral meshes including layered-earth geometries with topography or bathymetry and anomalous bodies which are allowed to be connected or to reach the surface.

@@

The 3D CSEM code \emg3d is a multigrid solver \citep{CMMP.64.Fedorenko} for electromagnetic diffusion following \cite{GP.06.Mulder}, with tri-axial electrical anisotropy, isotropic electric permittivity, and isotropic magnetic permeability. The matrix-free solver can be used as main solver or as preconditioner for one of the Krylov subspace methods implemented in SciPy, and the governing equations are discretized on a staggered grid by the finite-integration technique \citep{AEU.77.Weiland}, which is a finite-volume generalisation of a Yee grid. The code is written completely in Python using the NumPy and SciPy stacks, where the most time- and memory-consuming parts are sped up through jitted (just-in-time compiled) functions using Numba \citep{LLVM.15.Lam}. The strength of \emg3d is the matrix-free multigrid implementation, which is characterised by almost linear scaling both in terms of runtime (CPU) and memory (RAM) usage, and it is therefore a comparably low-memory consumption solver.

@@

Current development is focused on adding basic inversion capabilities, and a further plan is to implement a wrapper in order to use \emg3d as a frequency-domain CSEM solver within the grander \simpeg framework.

@@

\petgem permits to locate the source and receivers anywhere in the computational domain (e.g., sediments, seafloor, sea, ground), allowing to analyse the physical environment of the electric responses and how parameters impact them (e.g., frequency, conductivity, dependence on mesh setup, basis order, solver type, among others). Furthermore, \petgem implements a semi-adaptive mesh strategy ($hp$ mesh refinement) based on physical parameters and on polynomial order to satisfy quality criteria chosen by the user. Nonetheless, only Horizontal Electric Dipole (HED) has been implemented.

For the parallel forward modelling computations, a highly scalable MPI domain decomposition allows reducing runtimes and the solution of large-scale modelling cases. This strategy is capable of exploiting the parallelism offered by both modest multi-core computers and cutting-edge clusters (e.g., High-Performance Computing architectures).

@@

\simpeg is a modular toolbox for simulations and gradient based inversions in geophysics. Current functionality includes modelling and inversion capabilities for gravity, magnetics, direct current resistivity, induced polarisation, electromagnetics (time and frequency, controlled and natural sources such as magnetotellurics), and fluid flow (Richards equation). It is a community driven project that aims to support researchers by providing a flexible, extensible framework and to facilitate integration of data sets from a variety of geophysical methods, including joint inversions \citep{GJI.20.Astic}, by providing a common interface to each method. \mycom[DW]{Lindsey: «support researchers» downplays the fact that \simpeg is used by various companies as well\dots}

@@

The electromagnetic implementations use a staggered grid finite volume approach where physical properties are discretized at cell centres, fields on cell edges and fluxes on cell faces \citep{IEEE.66.Yee}.
%(for an overview of discretizing Maxwell's equations, we recommend \cite{B.SIAM.14.Haber}).
There are two different discretization strategies implemented for Maxwell's equations: (I) the EB-formulation, which discretizes the electric field ($\vec{e}$) on cell edges and the magnetic flux density ($\vec{b}$) on cell faces, and (II) the HJ-formulation, which discretizes the magnetic field ($\vec{h}$) on cell edges and the current density ($\vec{j}$) on cell faces. The physical properties electrical conductivity / resistivity and magnetic permeability are discretized in both cases at cell centres. Having multiple implementations allows for testing that compares results from each approach, as well as the representation of both electrical and magnetic sources on cylindrically symmetric meshes. \simpeg supports variable magnetic permeability and full-tensor anisotropy for the physical properties.

\simpeg interfaces to various solvers including \texttt{PARDISO} \citep{FGCS.04.Schenk}, \texttt{MUMPS} \citep{SIAM.01.Amestoy}, as well as those implemented in the \texttt{scipy.sparse.linalg} module \citep{NM.20.Virtanen}. For problems with many sources (e.g. airborne EM), \simpeg contains machinery for domain-decomposition approaches; see for example \cite{GEO.20.Fournier}. \mycom[DW]{only cite ones MUMPS etc, check and adjust}

\section{Numerical Verification}

@@

\itodo{ADD cross section plots!}

@@

\section{Acknowledgement}
